Tropopause Evolution in a Rapidly Intensifying Tropical Cyclone: A Static Stability Budget Analysis

### 2018-02-08 paper.tex
Added a lot to the Results and Discussion sections and changed Figures.

\paragraph{48-72 hours}
The third day of the simulation marked a dramatic change in the structure of the tropopause-layer static stability tendencies. During this time, static stability increased markedly in an upward-sloping region within the 30-60-km radial band (Fig.~\ref{fig:stab-48-72}a), and also increased within the 16.75-17.5-km layer out to at least the 200-km radius.
As this layer stabilized, the layer immediately below it destabilized in a broad region extending from 60-200 km.
Examination of the contribution from total advection (Fig.~\ref{fig:stab-48-72}b) reveals that advection no longer dominates the static stability tendencies.
Instead, a combination of vertical turbulence (Fig.~\ref{fig:stab-48-72}c) and radiation (Fig.~\ref{fig:stab-48-72}d) overcomes the destabilizing influence of advection to create the layer of increasing static stability.
Meanwhile, the destabilizing influence of vertical turbulence in a broad region below 17 km combines with a small region of destabilization due to radiation in the 50-120-km radial band combine to destabilize the layer below 16.5 km in the 50-200-km radial band.
Comparing the sum of advection and vertical turbulence (Fig.~\ref{fig:stab-48-72}e) to the sum of advection, vertical turbulence, and radiation (Fig.~\ref{fig:stab-48-72}f) reveals that radiation plays a fundamental role in the re-strengthening of the lower-stratospheric stable layer during this time.

\paragraph{72-96 hours}

@@

@@

%FIGURE 8%
\begin{figure*}[ht]
\centerline{\includegraphics[width=39pc]{figures/u.png}}
\caption{Radial velocity (m s\textsuperscript{-1}; filled contours), potential temperature (K; thick black contours), and cold point tropopause height (orange lines) averaged over (a) 0-24 hours, (b) 24-48 hours, and (c) 48-72 hours.}
\label{fig:u}
\end{figure*}

%FIGURE 9%
\begin{figure*}[ht]
\centerline{\includegraphics[width=39pc]{figures/w.png}}
\caption{Vertical velocity (cm s\textsuperscript{-1}; filled contours), potential temperature (K; thick black contours), and cold point tropopause height (orange lines) averaged over (a) 0-24 hours, (b) 24-48 hours, and (c) 48-72 hours.}
\label{fig:w}
\end{figure*}

%FIGURE 10%
\begin{figure*}[ht]
\centerline{\includegraphics[width=39pc]{figures/qtot.png}}
\caption{Total condensate mixing ratio (g kg\textsuperscript{-1}) and cold point tropopause height (orange lines) averaged over (a) 0-24 hours, (b) 24-48 hours, and (c) 48-72 hours.}
\label{fig:qtot}
\end{figure*}

%FIGURE 11%
\begin{figure*}[ht]
\centerline{\includegraphics[width=39pc]{figures/radten.png}}
\caption{Radiative heating rate (K hr\textsuperscript{-1}) and cold point tropopause height (orange lines) averaged over (a) 0-24 hours, (b) 24-48 hours, and (c) 48-72 hours.}
\label{fig:radten}
\end{figure*}

%FIGURE 12%
\begin{figure*}[ht]
\centerline{\includegraphics[width=39pc]{figures/khvten.png}}
\caption{Vertical eddy diffusivity (m\textsuperscript{2} s\textsuperscript{-2}; filled contours), cold point tropopause height (cyan lines), and radial velocity (m s\textsuperscript{-1}; thick black lines) averaged over (a) 0-24 hours, (b) 24-48 hours, and (c) 48-72 hours.}
\label{fig:radten}
\end{figure*}

>>>>

\paragraph{48-72 hours}
After the storm's maximum wind speed leveled off near 80 m s\textsuperscript{-1}, the magnitude of the static stability tendencies within the eye decreased to near zero (Fig.~\ref{fig:stab-48-72}a).

Outside of the eye, however, $N^2$ continued to increase just above the tropopause and decrease just below.
The sum of advection and vertical turbulence (Fig.~\ref{fig:stab-48-72}e) indicates that the increase of $N^2$ observed in the 17-18-km layer and inside of the 80-km radius cannot be attributed to these processes, since the sum of these two terms provided forcing for destabilization.
Instead, radiation (Fig.~\ref{fig:stab-48-72}d), provided the forcing for stabilization in this region.
Outside of the 80-km radius, both advection (Fig.~\ref{fig:stab-48-72}b) and vertical turbulence (Fig.~\ref{fig:stab-48-72}c) provided forcing for stabilization near the 18-km level.
The sum of the two terms indicates increasing $N^2$ near the 18-km level everywhere outside of the 80-km radius, but this stabilization is slightly weaker in the 90-120-km radial band than the observed value.
The addition of radiation (Fig.~\ref{fig:stab-48-72}f) provides the extra forcing for stabilization required to account for the observed increase in $N^2$.
Outside of the 120-km radius, the region of radiative forcing for stabilization slopes downward, and the increase in $N^2$ observed near 18 km can be explained entirely by a combination of advection and vertical turbulence.
The layer of decreasing $N^@$ observed near 17 km was forced primarily by vertical turbulence and radiation.
Within most of this region, advection provided strong forcing for stabilization, but this forcing was outweighed by the negative $N^@$ tendencies induced by a combination of vertical turbulence and radiation.

@@

  \subsection{The role of advection}

Advection played an important role in the tropopause-layer $N^2$ evolution at all stages of intensification.
To investigate the advective processes more closely, the individual contributions of horizontal and vertical advection are shown for all three time periods in Figs. \ref{fig:adv-00-24}-\ref{fig:adv-48-72}, along with the corresponding time-mean velocities and $\theta$.
In each of the three time periods, the $N^2$ tendencies due to the two advective components exhibit strong cancellation, consistent with flow that is nearly isentropic.
There are, however, some regions in which flow crosses $\theta$ surfaces; this flow accounts for all non-zero $N^2$ tendencies due to advection in the budget analyses.

  \paragraph{0-24 hours}
During the first day of the simulation, horizontal advection (Fig. \ref{fig:adv-00-24}a) generally forced a decrease in $N^2$ below 17 km and an increase in $N^2$ above.
The mean radial velocity during this period (Fig. \ref{fig:adv-00-24}c) reveals a region of outflow at and below 17 km and near-zero radial velocity above, with a layer of strong vertical wind shear straddling the tropopause.
The slight downward slope with radius of the $\theta$ surfaces near the tropopause (meaning that $\theta$ increases with radius) implies that the outflow in this layer locally acted to decrease $\theta$.
This negative $\theta$ tendency must decrease with height within the region of strong vertical wind shear and approach zero as the radial velocity diminishes to zero near 17 km.
These tendencies thus yielded an increase in the vertical gradient of $\theta$ above 17 km, which implies an increase in $N^2$, as was seen in Fig. \ref{fig:adv-00-24}a.
Note that although the radial gradient of $\theta$ was stronger in the 15-16-km layer than it was closer to the tropopause, the vertical shear of the radial wind was much smaller there, yielding a smaller $N^2$ tendency due to horizontal advection.

Notably, there existed strong $N^2$ tendencies due to horizontal advection inside of the 30-km radius even though the mean radial velocity in that region was near zero.
Close examination of the radial velocity at each output time step reveals the presence of inward-propagating waves in the lower stratosphere.
These waves amplified upon reaching r=0 and forced strong but highly transient $\theta$ tendencies due to advective processes.
Thus, although the radial velocity perturbations associated with these waves averaged out to zero, the non-linear effect of these waves on the $N^2$ field remained.

The $N^2$ tendencies forced by horizontal advection were completely canceled out by vertical advective tendencies (Fig. \ref{fig:adv-00-24}b).
As was seen in Fig. \ref{fig:stab-00-24}b, the sum of the two terms yielded a layer of increasing $N^2$ below 17 km and decreasing $N^2$ above, meaning that vertical advective processes dominated the $N^2$ tendency during the initial spin-up period.

@@

%FIGURE 8%
\begin{figure*}[ht]
\centerline{\includegraphics[width=39pc]{figures/h000-h024-adv.png}}
\caption{The contribution to the change in $N^2$ over the 0-24-hour period (10\textsuperscript{-4} s\textsuperscript{-2} (24 hr)\textsuperscript{-1}) by (a) horizontal advection and (b) vertical advection. (c) The radial velocity (m s\textsuperscript{-1}; filled contours), potential temperature (K; thick black contours), and cold-point tropopause height (orange line) averaged over the 0-24-hour period. (d) The vertical velocity (cm s\textsuperscript{-1}; filled contours), potential temperature (K; thick black contours), and cold-point tropopause height (orange line) averaged over the 0-24-hour period.}
\label{fig:adv-00-24}
\end{figure*}

%FIGURE 8%
%\begin{figure*}[ht]
%\centerline{\includegraphics[width=39pc]{figures/u.png}}
%\caption{Radial velocity (m s\textsuperscript{-1}; filled contours), potential temperature (K; thick black contours), and cold point tropopause height (orange lines) averaged over (a) 0-24 hours, (b) 24-48 hours, and (c) 48-72 hours.}
%\label{fig:u}
%\end{figure*}

%FIGURE 9%
\begin{figure*}[ht]
\centerline{\includegraphics[width=39pc]{figures/h024-h048-adv.png}}
\caption{As in Fig. \ref{fig:adv-00-24}, but for the 24-48-hour period.}
\label{fig:adv-24-48}
\end{figure*}

%FIGURE 9%
%\begin{figure*}[ht]
%\centerline{\includegraphics[width=39pc]{figures/w.png}}
%\caption{Vertical velocity (cm s\textsuperscript{-1}; filled contours), potential temperature (K; thick black contours), and cold point tropopause height (orange lines) averaged over (a) 0-24 hours, (b) 24-48 hours, and (c) 48-72 hours.}
%\label{fig:w}
%\end{figure*}

%FIGURE 10%
\begin{figure*}[ht]
\centerline{\includegraphics[width=39pc]{figures/h048-h072-adv.png}}
\caption{As in Fig. \ref{fig:adv-00-24}, but for the 48-72-hour period.}
\label{fig:adv-48-72}
\end{figure*}

%FIGURE 11%
\begin{figure*}[ht]
\centerline{\includegraphics[width=39pc]{figures/qtot.png}}
\caption{Total condensate mixing ratio (g kg\textsuperscript{-1}) and cold point tropopause height (orange lines) averaged over (a) 0-24 hours, (b) 24-48 hours, and (c) 48-72 hours.}
\label{fig:qtot}
\end{figure*}

%FIGURE 12%
\begin{figure*}[ht]
\centerline{\includegraphics[width=39pc]{figures/radten.png}}
\caption{Radiative heating rate (K hr\textsuperscript{-1}) and cold point tropopause height (orange lines) averaged over (a) 0-24 hours, (b) 24-48 hours, and (c) 48-72 hours.}
\label{fig:radten}
\end{figure*}

%FIGURE 13%
\begin{figure*}[ht]
\centerline{\includegraphics[width=39pc]{figures/khvten.png}}
\caption{Vertical eddy diffusivity (m\textsuperscript{2} s\textsuperscript{-2}; filled contours), cold point tropopause height (cyan lines), and radial velocity (m s\textsuperscript{-1}; thick black lines) averaged over (a) 0-24 hours, (b) 24-48 hours, and (c) 48-72 hours.}
\label{fig:diff}
\end{figure*}
